[Re] Marriage and happiness: Providing evidence against a relationship between inequality and happiness in Oishi, Kesebir, and Diener (2011)

### 2020-02-23 analysis/paper.Rmd
Adds details for table 1 and table 2 replication process, also adds details to conclusion/intro

The dataset from gssr contained a larger time interval than the original paper, it was filtered back to match the original intervals. This dataset was also merged with the original US Census Bureau data found in Gini_families.csv[@data] by appending the total column to the GSS data using year as the key. The "happy" variable was re-encoded so that the numerical factor would reflect an increasing level of happiness as per the original code.

@@

@@

@@

The R function cor.test() was used to get the Pearson correlation coefficient between the different independent variables (income, age, sex, white, married) and the response variable, happiness. Grundberg et al.'s code was slightly modified to keep the same encoding of variables given the new dataset input.

@@

The  to almost exactly reproduce the results from Table 1 in the original paper that shown in Figure \@ref(fig:table1). The only difference is the estimate for Sex was 0.01 in the publication compared to 0.00 in our replication. This may be due to rounding versions or an underlying exclusion of outliers we saw with the above plot.

@@

Again we were able to almost identically reproduce the results from Table 2 in the original paper, shown in Figure \@ref(fig:table2). Aside from slight variability in the second and third decimal places, our results from the linear models are very close. There are several differences wit the number of observations, listed in the bottom of the table, which is to be expected as we are using a package with the GSS data and it is likely not an identical snapshot to the one used in the original paper. Overall though these two tables were indeed highly reproducible even with different input data.

@@

Replication of this paper was ultimately contingent on tracing back the original dataset. Without this resource being cited and publicly available, it would not have been possible to replicate the different tables and figure presented in the original paper.

>>>>

The dataset from gssr contained a larger time interval than the original paper, it was filtered back to match the original intervals. Only years prior to 2008 are used in both analyses. This dataset was also merged with the original US Census Bureau data found in Gini_families.csv[@data] by appending the total column to the GSS data using year as the key.

Several encodings for the data were updated in the original paper which were replicated in order to allow accurate comparisons. The "happy" variable was re-encoded in a reversed scheme so that the numerical factor would reflect an increasing level of happiness. The income data was re-encoded on a log scale. Race was encoded numerically where "White" was set to 1. The "married" variable was also set to numeric encodings with married respondents denoted by those answering 1.

## Alterations

Aside from gathering the input data from a new source, several different packages were used in addition to ensure the reproduced results were adequately tested. These will be discussed further in the respective sections below.

@@

Figure 3 in the original publication plots mean happiness of respondentsover time. This data is categorized into 3 groups by racial identity: White, Black and Other. Best-fitting regression lines are included to display the trend of mean happiness for the 3 populations.

@@

Table 1 in the original publication explores the correlations between happiness and various control variables found in the dataset.

@@

The R function cor.test() was used to get the Pearson correlation coefficient between the different independent variables (income on a log scale, age, sex, white, married) and the response variable, happiness. Grundberg et al.'s code was slightly modified to keep the same encoding of variables given the new dataset input.

@@

```{r, table1-replication, fig.cap="Replication of Table 1", echo=FALSE}
knitr::include_graphics("table_1.png")
```

We were able to almost exactly reproduce the results from Table 1 in the original paper that shown in Figure \@ref(fig:table1). The only difference is the estimate for Sex was 0.01 in the publication compared to 0.00 in our replication. This may be due to rounding versions or an underlying exclusion of outliers we saw with the above plot.

The steps taken to reproduce this table are identical to those in the original publication aside from a new input dataset and the use of a combination of the kableExtra and xtable R packages to produce the output compared with xtable alone used in the original publication.

Table 2 in the original publication explores the output of various linear models on the dependent variable, happiness. This is central to the message of the paper and was therefore critical to investigate.

@@

```{r, table2-replication, fig.cap="Replication of Table 2", echo=FALSE}
knitr::include_graphics("table_2.png")
```

Again we were able to almost identically reproduce the results from Table 2 in the original paper, shown in Figure \@ref(fig:table2). Aside from slight variability in the second and third decimal places, our results from the linear models are very close. There are several differences with the number of observations, listed in the bottom of the table, which is to be expected as we are using a package with the GSS data and it is likely not an identical snapshot to the one used in the original paper. However, this variability is minor in scale.Even though slight variabilities were observed in different cells throughout this table, no disparities were large enough to impact the significance of the results as denoted in the tables with asterisks. Therefore, even with new data sources we found that the conclusions hold for the linear model analysis.

Again, the steps taken to reproduce this table were nearly identical as the original paper aside from a new dataset source and the use of the kable and stargazer packages to produce the output as opposed to strictly stargazer LaTeX table results.

@@

Replication of this paper was ultimately contingent on tracing back the original dataset. Without this resource being cited and publicly available, it would not have been possible to replicate the different tables and figure presented in the original paper. The lack of necessary information included in the published data provided for replication is something that should likely be addressed by the authors as this was a major deterrent to the goal of reproducibility.

Under conditions of both a new source dataset and different R packages used for data manipulation and visualization, the 3 exhibits in the original publication were able to be replicated. With only minor discrepancies in the numbers themselves, the conclusions drawn from the paper's graphics were not significantly altered by any means. Successful replication of this paper is largely a result of transparency in the original publication providing clarity around data encoding and the models used for analysis. This provided a platform upon which we were able to adapt the code and generate results that in large part align directly with those in the publication.
